\documentclass[acmsmall,nonacm,screen,10pt]{acmart}
\usepackage{preamble}

%% ── Document metadata ─────────────────────────────────────

\title{Swift Concurrency as a Capability-Region Type System}
\subtitle{Capability, region, and affine transfer in Swift 6.2}

\author{Yasuhiro Inami (@inamiy)}
\thanks{This paper was produced with the assistance of Claude Opus 4.6 and GPT-5.4.}
\affiliation{%
  \institution{try! Swift Tokyo 2026}
  \country{Japan}
}
\date{Apr 13, 2026}

\authorsaddresses{}

\begin{document}

\begin{abstract}
Swift Concurrency exposes programmers to individually sensible rules whose combined behavior is still difficult to predict from surface syntax alone. The hardest cases mix actor isolation, region-based isolation, \code{@Sendable}, \code{sending}, nonisolated async execution, and dynamic \code{@isolated(any)} function values. This paper presents a rule-level account of Swift~6.2 concurrency based on three ingredients: execution capability, value region, and affine environment update. Capability answers \emph{where} a computation runs. Region answers \emph{where} a non-\code{Sendable} value currently lives. Affine update explains \emph{why} a call may preserve access, refine a binding, or consume a value from the caller-side environment. The resulting account clarifies five recurring phenomena: same-isolation \code{sending} does not necessarily consume, ordinary same-isolation calls bind disconnected values into actor regions, cross-isolation results require an explicit \code{sending} asymmetry, \code{@isolated(any)} behaves as a dynamic isolation interface rather than an ordinary actor annotation, and \code{@nonisolated async} after SE-0461 inherits caller execution without collapsing into \code{@concurrent}~\cite{se0461}. The paper validates this model against a focused compiler probe suite covering conversion behavior, task capture, actor-instance closure inference, dynamic isolation identity, and result-transfer corner cases. The appendix prints the complete rule inventory so the paper serves as both an argument and a reference.
\end{abstract}

\keywords{Swift Concurrency, actor isolation, region-based isolation, \texttt{Sendable}, \texttt{sending}, \texttt{@isolated(any)}, type systems.}

\maketitle
{\small Last Update: April 13, 2026}
\bigskip

\input{sections/introduction}
\input{sections/background}
\input{sections/model}
\input{sections/conversion-rules}
\input{sections/kernel-rules}
\input{sections/case-studies}
\input{sections/validation}
\input{sections/discussion}
\input{sections/conclusion}

\appendix
\clearpage
\input{sections/appendix-notation}
\clearpage
\input{sections/appendix-relation-and-boundary-rules}
\clearpage
\input{sections/appendix-full-typing-rules}
\clearpage
\input{sections/appendix-examples}

\clearpage
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}
